The Essence of Summarization\\ in Pushdown Flow Analyses\\ \textbf{Supplemental Material}

### 2014-02-19 content.tex
minor paper fixes. Switched to sigplan.

Programs in higher-order languages heavily use function calls and method dispatch as part of their control flow.
%
Until recently, flow analyses for higher-order languages could not handle return flow precisely \citep{ianjohnson:vardoulakis-lmcs11, dvanhorn:Earl2010Pushdown}, which leads to several spurious paths (and thus false positives) due to the pervasiveness of function or method calls and subsequent returns.
%
These works, called CFA2 and PDCFA respectively, use pushdown automata as their approximation's target model of computation.
%
They are hence called ``pushdown analyses.''%
%
\footnote{We refer to finite model analyses as ``regular analyses'' after regular languages.}
%
CFA2 and PDCFA have difficult details to easily apply to an off-the-shelf semantics---especially if they feature non-local control transfer that breaks the pushdown model.
%%

%%
There is a systematic process for transforming off-the-shelf programming language semantics into a form amenable to \emph{regular} analysis that has been widely applied with great success to production programming languages (a technique called abstracting abstract machines, or AAM~\citep{dvanhorn:VanHorn2010Abstracting}).
%
A contribution of this paper is a systematic process to construct \emph{pushdown} analyses of programming languages, due to the precision (and often performance) benefits.
%

@@

%%
Testing new ideas in analysis for improving precision or increasing performance can be a difficult venture.
%
We contend that the machinery that we employ in the abstract should have a concrete counterpart that maintains the meaning of the language so that we can see their effect without introducing approximations which might mask correctness issues.
%
Abstraction should be a simple process that is ``obviously correct.''
%
The framework we present in this paper is applicable in the concrete such that a point-wise abstraction leads to the pushdown analyses in the literature.
%%

@@

%%
Continuations count as ``extra'' in this semantics since functions cannot inspect the stack, so $\krt{\mctx}$ might be considered a special allocation strategy for continuations in the AAM viewpoint.
%
We separate the table of continuations, $\mktab$, from the store, $\mstore$, since $\rt$ continuations contain stores---this leads to a recursive data-structure that we are trying to avoid.
%
The store is an important ingredient to maintaining enough precision to keep a pushdown abstraction.
%
The semantic spaces for the machine are modified thusly:

@@

\paragraph{The role of summaries:} notice the additional $\Memo$ component.
%
A ``summary edge'' in CFA2 or equivalently an $\epsilon$-edge in PDCFA is an edge from the source of a push edge to the target of the matching pop edge.
%
``Matching'' here means there is a path through machine reductions that don't change the stack, or through summary edges.
%
In our model, we only ``push'' when we call a function, and we only ``pop'' when we return with a value.
%
There is an analogy to something more operational: summary edges embody memoization.
%
Instead of following an entire path through a call to return a value, we simply jump from the call to the return with the result of the call, with the resulting context changes---in this case only the store \textsc{[memo lookup]}.
%%

\begin{figure}
  \centering
  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
  \begin{tabular}{r|ll}
    \hline
% variable lookup
    $\tpl{(\svar\mvar, \menv), \mstore, \mkont, \mktab, \mmemo}$
    &
    $\tpl{\mval,\mstore,\mkont, \mktab, \mmemo}$ if $\mval \in \mstore(\menv(\mvar))$
    &
    \textsc{[variable lookup]}
    \\
% application
    $\tpl{(\sapp{\mexpri0}{\mexpri1}, \menv), \mstore, \mkont, \mktab, \mmemo}$
    &
    $\tpl{(\mexpri0, \menv), \mstore, \kcons{\kar{\mexpri1,\menv}}{\mkont}, \mktab, \mmemo}$
    &
    \textsc{[function eval]}
    \\
% argument evaluation
    $\tpl{\mval, \mstore, \kcons{\kar{\mexpr,\menv}}{\mkont}, \mktab, \mmemo}$
    &
    $\tpl{(\mexpr, \menv), \mstore, \kcons{\kfn{\mval}}{\mkont}, \mktab, \mmemo}$
    &
    \textsc{[argument eval]}
    \\
% function call
    $\tpl{\mval,\mstore,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont},\mktab,\mmemo}$
    & % actually do call,
    $\tpl{\mpoint,
          \mstore',
          \krt{\mpoint, \mstore'},
          \mktab',
          \mmemo}$
    &
    \textsc{[un-memoized call]}
     \\ & \quad if $(\mpoint, \mstore') \notin \dom(\mmemo)$
\\
    & or \\
    & $\tpl{\mval_\mathit{result},
            \mstore'',
            \mkont,
            \mktab',
            \mmemo}$
    & \textsc{[memo lookup]}
    \\ & \quad if $(\mval_\mathit{result},\mstore'') \in \mmemo(\mpoint,\mstore')$
    \\ % or, lookup in table
%    & $\tpl{\mval, \mstore', \mkont,\mktab,\mmemo}$ if $\mval \in \mmemo(\mpoint, \mstore')$ \\    
    where & $\mpoint = (\mexpr, \extm{\menv}{\mvar}{\maddr})$ \\
          & $\mstore' = \joinone{\mstore}{\maddr}{\mval}$ \\
          & $\mktab' = \joinone{\mktab}{(\mpoint, \mstore')}{\mkont}$
    \\
% return
    $\tpl{\mval, \mstore, \krt{\mpoint,\mstore'}, \mktab, \mmemo}$
    &
    $\tpl{\mval, \mstore, \mkont, \mktab, \joinone{\mmemo}{(\mpoint, \mstore')}{(\mval,\mstore)}}$
    &
    \textsc{[memoize/return]}
    \\ & \quad if $\mkont \in \mktab(\mpoint, \mstore')$
  \end{tabular}
  \caption{The summarizing tabular stack machine}
  \label{fig:summary-semantics}
\end{figure}

@@

%%
We should expect that a pushdown analysis would predict this evaluates to true, and there are no loops in the program.
%
0-CFA~\citep{dvanhorn:Shivers:1991:CFA} claims there is a loop from the second call of \texttt{id} to the first, and thusly predicts this program evaluates to true or false.
%
PDCFA claims there is no loop and that the result is true.
%
The example evaluation is in \autoref{fig:ex-eval}.
%
Here $\mkont_0 = \mathbf{lt}({\tt y}, {\tt (let\ ([z\ (id\ 1)])\ (}\le\ {\tt y}\ {\tt z))}, \iota)$.
%
We maintain enough context to distinguish the return points of \texttt{id} to not rebind \texttt{y} to 1.
%
When determining control flow through the expression, we consult $\mktab$ to continue past function returns, so there is no confusion about a back edge from the second call to \texttt{id}.
%

@@

Abstract garbage collection~\citep{dvanhorn:Might:2006:GammaCFA} is a powerful technique for regaining precision in an abstract interpretation.
%
The followup to PDCFA~\citep{dvanhorn:Earl2012Introspective} that allows inspecting the stack in order to perform sound GC introduces an entirely new class of automata for this one purpose.
%
In our view with tabular semantics, the GC story mimics the simplicity of GC in the regular setting, without the detour into automata theory.
%
We follow the GC style of \citet{dvanhorn:Might:2006:GammaCFA} with a stop-and-copy collector.
%
Address reachability is in terms of a base ``touches'' function, $\touches$.
%
Entries in $\mktab$ might be circular, so we compute the continuation frames in a fixed-point before computing touched addresses:
\begin{center}
  \,
  \begin{minipage}{0.32\linewidth}
    $\touchesk{\mkont}(\mktab) = \setbuild{\maddr}{\mkont \rightsquigarrow_\mktab^*\maddr}$
  \end{minipage}
  \begin{minipage}{0.36\linewidth}
    $\touchesf(\kar{\mexpr,\menv}) = \mathit{rng}(\menv|_{\mathit{fv}(\mexpr)})$
  \end{minipage}
  \begin{minipage}{0.3\linewidth}
    $\touchesf(\kfn{\mval}) = \touches(\mval)$
  \end{minipage}
\end{center}
\begin{mathpar}
  \inferrule{ }{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab \mkont}
  \quad
  \inferrule{\maddr \in \touchesf(\mkframe)}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab \maddr}
  \quad
  \inferrule{\mkont \rightsquigarrow_\mktab \mkont'}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab \mkont'}
  \quad
  \inferrule{\mkont \rightsquigarrow_\mktab \maddr}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab \maddr}
%  \inferrule{\mkont \rightsquigarrow_\mktab (\mkont' \text{ or } \maddr)}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab (\mkont' \text{ or } \maddr)}
  \quad
  \inferrule{\mkont \in \mktab(\mctx)}{\krt{\mctx} \rightsquigarrow_\mktab \mkont}
\end{mathpar}

Garbage collection ($\Gamma$) restricts the heap to just the reachable addresses, so a semantics with GC introduces the following reduction rule:
\begin{center}
  \begin{minipage}{0.3\linewidth}
    \begin{align*}
      \frac{\mstate \stepto \mstate'}{\mstate \stepto \Gamma(\mstate')}
    \end{align*}
  \end{minipage}
  \begin{minipage}{0.65\linewidth}
    \begin{align*}
      \Gamma(\tpl{(\mexpr,\menv),\mstore,\mkont, \mktab, \mmemo}) &= \tpl{(\mexpr,\menv),\mstore|_{\reaches(\mathit{root}, \mstore)}, \mkont, \mktab, \mmemo} \\
      \text{where } \mathit{root} &= \touchesk{\mkont}(\mktab) \cup
      \mathit{rng}(\menv|_{\mathit{fv}(\mexpr)})
    \end{align*}
  \end{minipage}
\end{center}
%
The ``reaches'' metafunction ($\reaches$) is a transitive closure of ``touches'' as it weaves through the store, starting from some root set of addresses:

@@

The creators of CFA2 had a clear goal of harnessing the extra information a pushdown model provides to produce a high-precision analysis that works well in practice.
%
This resulted in more than just the call/return matching of the previous section, which is why we are showing the two separately.
%

@@

%%
The second of these is to ameliorate a problem they call ``fake rebinding.''
%
Bindings in the abstract represent several values, but we don't want to reference a variable \texttt{x} in different places in a function and have it resolve to different values
\footnote{In AAM, a variable reference non-deterministically steps to all possible values associated with that variable.}.
%
If subsequent variable references are not the some, it looks as if the variable was rebound; it hasn't, and thus it is a ``fake rebinding.''
%
%CFA2 does not step to all values on variable reference, but instead carries all its values around in superposition until they need to be observed at, say, a function call.
%
%We give a simplified semantics that is more along the AAM style, but CFA2's approach can easily be recovered from it.\footnote{Our Redex model implements the fake rebinding strategy CFA2 itself employs.}
%%

@@

A tabular semantics for the CES$\xi$K machine follows the previous section, except that $\mktab$ now stores $(\mkont, \mframe)$ pairs, and returns reinstate the saved $\mframe$:
\begin{align*}
  \tpl{\mval,\mstore,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont},\mktab,\mmemo} &\stepto
  \tpl{\mpoint,
    \mstore',
    \mframe',
    \krt{\mpoint, \mstore'},
    \mktab',
    \mmemo} \\
  \text{where }
  & \mpoint = (\mexpr, \extm{\menv}{\mvar}{\maddr}) \\
  & (\mstore',\mframe') = \bind(\mstore,\maddr,\mvar,\mval) \\
  & \mktab' = \joinone{\mktab}{(\mpoint, \mstore')}{(\mkont,\mframe)}
\\[2pt]
 \tpl{\mval, \mstore, \mframe, \krt{\mpoint,\mstore'}, \mktab, \mmemo} &\stepto
 \tpl{\mval, \mstore, \mframe', \mkont, \mktab, \joinone{\mmemo}{(\mpoint, \mstore')}{(\mval,\mstore)}}
  \\ & \quad\text{if } (\mkont,\mframe') \in \mktab(\mpoint, \mstore')
\end{align*}

@@

\begin{center}
  \begin{minipage}{0.45\linewidth}
    \begin{align*}
      \mstate \in \State &::= \tpl{\mpoint, \mstore, \mkont, \mmkont} \\
      \mpoint \in \Point &::= (\mexpr, \menv) \alt \mval
    \end{align*}
  \end{minipage}
  \begin{minipage}{0.50\linewidth}
    \begin{align*}
      \mval \in \Value &::= \vclo{\slam{\mvar}{\mexpr}}{\menv} \alt \vcomp{\mkont} \\
      \mmkont \in \MKont &::= \kmt \alt \mkapp{\mkont}{\mmkont}
    \end{align*}
  \end{minipage}
\end{center}
\begin{figure}
  \centering
  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
  \begin{tabular}{r|ll}
    \hline
% Reset
    $\tpl{(\sreset{\mexpr}, \menv), \mstore, \mkont, \mmkont}$
    &
    $\tpl{(\mexpr, \menv), \mstore, \kmt, \mkapp{\mkont}{\mmkont}}$
    &
    \textsc{[push prompt]}
    \\
% Pop prompt
    $\tpl{\mval, \mstore, \kmt, \mkapp{\mkont}{\mmkont}}$
    &
    $\tpl{\mval, \mstore, {\mkont}, {\mmkont}}$
    &
    \textsc{[pop prompt]}
    \\
% Shift
    $\tpl{(\sshift{\mvar}{\mexpr}, \menv), \mstore, \mkont, \mmkont}$
    &
    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \mstore',\kmt,\mmkont}$
    &
    \textsc{[capture continuation]}
    \\ & where $\mstore' = \joinone{\mstore}{\maddr}{\vcomp{\mkont}}$
    \\
% continuation call
    $\tpl{\mval,\mstore,\kcons{\kfn{\vcomp{\mkont'}}}{\mkont}, \mmkont}$
    &
    $\tpl{\mval, \mstore, \mkont', \mkapp{\mkont}{\mmkont}}$
    &
    \textsc{[compose continuation]}
  \end{tabular}  
  \caption{Machine semantics for shift/reset}
  \label{fig:shift-reset}
\end{figure}

@@

\begin{figure*}
  \centering
  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
  \begin{tabular}{r|l}
    \hline
% Reset
    $\tpl{(\sreset{\mexpr}, \menv), \mstore, \mkont, \mmkont, \mktab, \mmemo}$
    &
    $\tpl{\mpoint, \mstore, \kmt, \kprompt{\mctx}, \joinone{\mktab}{\mctx}{(\mkont,\mmkont)}, \mmemo}$
    \\
    where & $\mpoint = (\mexpr, \menv)$, $\mctx = (\mpoint, \mstore)$.
    \\
% Pop prompt
    $\tpl{\mval, \mstore, \kmt, \kprompt{\mctx}, \mktab, \mmemo}$
    &
    $\tpl{\mval, \mstore, {\mkont}, {\mmkont}, \mktab, \joinone{\mmemo}{\mctx}{(\mval,\mstore)}}$
    if $(\mkont,\mmkont) \in \mktab(\mctx)$
    \\
% Shift
    $\tpl{(\sshift{\mvar}{\mexpr}, \menv), \mstore, \mkont, \mmkont, \mktab, \mmemo}$
    &
    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \joinone{\mstore}{\maddr}{\vcomp{\mkont}},\kmt,\mmkont,\mktab,\mmemo}$
    \\
% continuation call
    $\tpl{\mval,\mstore,\kfn{\vcomp{\mkont'}, \mkont}, \mmkont, \mktab, \mmemo}$
    &
    $\tpl{\mval, \mstore, \mkont', \kprompt{\mctx}, \mktab',\mmemo}$
    \\ & \quad if $\mctx \notin \dom(\mmemo)$
    \\
    & or \\
    & $\tpl{\mval', \mstore', \mkont, \mmkont, \mktab',\mmemo}$
    \\ & \quad if $(\mval',\mstore') \in \mmemo(\mctx)$
    \\
    where & $\mctx = (\vcomp{\mkont'}, \mval, \mstore)$ \\
          & $\mktab' = \joinone{\mktab}{\mctx}{(\mkont,\mmkont)}$
    \\
% Memoized continuation call
% continuation call
    $\tpl{\mval,\mstore,\kcons{\kfn{\vcomp{\mkont'}}}{\mkont}, \mmkont, \mktab, \mmemo}$
    &
    $\tpl{\mval', \mstore', \mkont, \mmkont, \mktab,\mmemo}$ if $(\mval',\mstore') \in \mmemo(\mkont',\mval,\mstore)$
  \end{tabular}  
  \caption{Faulty table-based semantics for shift/reset}
  \label{fig:shift-reset-table0}
\end{figure*}

>>>>

Programs in higher-order languages heavily use function calls and method dispatch for control flow.
%
Standard flow analyses' imprecise handling of returns damages all specific analyses' precision.
%
Recent techniques match calls and returns precisely \citep{ianjohnson:vardoulakis-lmcs11, dvanhorn:Earl2010Pushdown} and build smaller models more quickly than a standard 0-CFA (evaluation predicted 2-5 times more constant bindings).
%
These works, called CFA2 and PDCFA respectively, use pushdown automata as their approximation's target model of computation.
%
They are hence called ``pushdown analyses.''%
%
\footnote{We refer to finite model analyses as ``regular analyses'' after the regular languages of traces they realize.}
%
CFA2 and PDCFA have difficult details to easily apply to an off-the-shelf semantics---especially if they feature non-local control transfer that breaks the pushdown model.
%%

%%
There is a systematic process for transforming off-the-shelf programming language semantics into a form amenable to \emph{regular} analysis that has been widely applied to production programming languages (a technique called abstracting abstract machines, or AAM~\citep{dvanhorn:VanHorn2010Abstracting}).
%
A contribution of this paper is a systematic process to construct \emph{pushdown} analyses of programming languages, due to the precision (and often performance) benefits.
%

@@

%%
Testing new ideas in analysis for improving precision or increasing performance can be a difficult venture.
%
Premature abstraction can mask correctness issues with complicated machinery we employ in our analyses\footnote{Oh so many personal experiences}.
%
We contend that analysis machinery should have an executable concrete counterpart that maintains the meaning of the language for effective testing.
%
Abstraction should then be a simple process that is ``obviously correct.''
%
The framework we present in this paper is applicable in the concrete such that a point-wise abstraction leads to the pushdown analyses in the literature.
%%

@@

%%
Continuations count as ``extra'' in this semantics since functions cannot inspect the stack, so $\krt{\mctx}$ might be considered a special allocation strategy for continuations in the AAM viewpoint.
%
We separate the table of continuations, $\mktab$, from the store, $\mstore$, since $\rt$ continuations contain stores---this leads to a recursive data-structure that we are trying to avoid.
%
The store is an important ingredient to maintaining enough precision to keep a pushdown abstraction.
%
The semantic spaces for the machine are modified as follows:

@@

\paragraph{The role of summaries:} notice the additional $\Memo$ component.
%
A ``summary edge'' in CFA2 or equivalently an $\epsilon$-edge in PDCFA is an edge from the source of a push edge to the target of the matching pop edge.
%
``Matching'' here means there is a path through machine reductions that don't change the stack, or through summary edges.
%
In our model, we only ``push'' when we call a function, and we only ``pop'' when we return with a value.
%
There is an analogy to something more operational: summary edges embody memoization.
%
Instead of following an entire path through a call to return a value, we simply jump from the call to the return with the result of the call, with the resulting context changes---in this case only the store (\todo{show in figure})
%%

\begin{figure}
  \centering
  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
  \begin{tabular}{r|l}%{r|ll}
    \hline
% variable lookup
    $\tpl{(\svar\mvar, \menv), \mstore, \mkont, \mktab, \mmemo}$
    &
    $\tpl{\mval,\mstore,\mkont, \mktab, \mmemo}$ if $\mval \in \mstore(\menv(\mvar))$
%    & \textsc{[variable lookup]}
    \\
% application
    $\tpl{(\sapp{\mexpri0}{\mexpri1}, \menv), \mstore, \mkont, \mktab, \mmemo}$
    &
    $\tpl{(\mexpri0, \menv), \mstore, \kcons{\kar{\mexpri1,\menv}}{\mkont}, \mktab, \mmemo}$
%    & \textsc{[function eval]}
    \\
% argument evaluation
    $\tpl{\mval, \mstore, \kcons{\kar{\mexpr,\menv}}{\mkont}, \mktab, \mmemo}$
    &
    $\tpl{(\mexpr, \menv), \mstore, \kcons{\kfn{\mval}}{\mkont}, \mktab, \mmemo}$
%    & \textsc{[argument eval]}
    \\
% function call
    $\tpl{\mval,\mstore,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont},\mktab,\mmemo}$
    & % actually do call,
    $\tpl{\mpoint,
          \mstore',
          \krt{\mpoint, \mstore'},
          \mktab',
          \mmemo}$
%    & \textsc{[un-memoized call]}
     \\ & \quad if $(\mpoint, \mstore') \notin \dom(\mmemo)$
\\
    & or \\
    & $\tpl{\mval_\mathit{result},
            \mstore'',
            \mkont,
            \mktab',
            \mmemo}$ % & \textsc{[memo lookup]}
    \\ & \quad if $(\mval_\mathit{result},\mstore'') \in \mmemo(\mpoint,\mstore')$
    \\ % or, lookup in table
%    & $\tpl{\mval, \mstore', \mkont,\mktab,\mmemo}$ if $\mval \in \mmemo(\mpoint, \mstore')$ \\    
    where & $\mpoint = (\mexpr, \extm{\menv}{\mvar}{\maddr})$ \\
          & $\mstore' = \joinone{\mstore}{\maddr}{\mval}$ \\
          & $\mktab' = \joinone{\mktab}{(\mpoint, \mstore')}{\mkont}$
    \\
% return
    $\tpl{\mval, \mstore, \krt{\mpoint,\mstore'}, \mktab, \mmemo}$
    &
    $\tpl{\mval, \mstore, \mkont, \mktab, \joinone{\mmemo}{(\mpoint, \mstore')}{(\mval,\mstore)}}$
%    & \textsc{[memoize/return]}
    \\ & \quad if $\mkont \in \mktab(\mpoint, \mstore')$
  \end{tabular}
  \caption{The summarizing tabular stack machine}
  \label{fig:summary-semantics}
\end{figure}

@@

%%
We should expect that a pushdown analysis would predict this evaluates to true, and there are no loops in the program.
%
0-CFA~\citep{dvanhorn:Shivers:1991:CFA} claims there is a loop from the second call of \texttt{id} to the first, and thus predicts this program evaluates to true or false.
%
PDCFA claims there is no loop and that the result is true.
%
The example evaluation is in \autoref{fig:ex-eval}.
%
Here $\mkont_0 = \mathbf{lt}({\tt y}, {\tt (let\ ([z\ (id\ 1)])\ (}\le\ {\tt y}\ {\tt z))}, \iota)$.
%
We maintain enough context to distinguish the return points of \texttt{id} to not rebind \texttt{y} to 1.
%
When determining control flow through the expression, we consult $\mktab$ to continue past function returns, so there is no confusion about a back edge from the second call to \texttt{id}.
%

@@

Abstract garbage collection~\citep{dvanhorn:Might:2006:GammaCFA} is a powerful technique for regaining precision in an abstract interpretation.
%
The followup to PDCFA~\citep{dvanhorn:Earl2012Introspective} that allows inspecting the stack in order to perform sound GC introduces an entirely new class of automata for this one purpose.
%
In our view with tabular semantics, the GC story mimics the simplicity of GC in the regular setting, without the detour into automata theory.
%
We follow the GC style of \citet{dvanhorn:Might:2006:GammaCFA} with a stop-and-copy collector.
%
Address reachability is in terms of a base ``touches'' function, $\touches$.
%
Entries in $\mktab$ might be circular, so we compute the continuation frames in a fixed-point before computing touched addresses:
\begin{center}
  \,
    $\touchesk{\mkont}(\mktab) = \setbuild{\maddr}{\mkont \rightsquigarrow_\mktab^*\maddr}$
\\[4pt]
  \begin{minipage}{0.5\linewidth}
    $\touchesf(\kar{\mexpr,\menv}) = \mathit{range}(\menv|_{\mathit{fv}(\mexpr)})$
  \end{minipage}
  \begin{minipage}{0.4\linewidth}
    $\touchesf(\kfn{\mval}) = \touches(\mval)$
  \end{minipage}
\end{center}
\begin{mathpar}
  \inferrule{ }{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab \mkont}
  \quad
  \inferrule{\maddr \in \touchesf(\mkframe)}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab \maddr}
  \quad
  \inferrule{\mkont \rightsquigarrow_\mktab \mkont'}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab \mkont'}
  \quad
  \inferrule{\mkont \rightsquigarrow_\mktab \maddr}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab \maddr}
%  \inferrule{\mkont \rightsquigarrow_\mktab (\mkont' \text{ or } \maddr)}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab (\mkont' \text{ or } \maddr)}
  \quad
  \inferrule{\mkont \in \mktab(\mctx)}{\krt{\mctx} \rightsquigarrow_\mktab \mkont}
\end{mathpar}

Garbage collection ($\Gamma$) restricts the heap to just the reachable addresses, so a semantics with GC introduces the following reduction rule:
\begin{center}
  \begin{minipage}{0.3\linewidth}
    \begin{align*}
      \frac{\mstate \stepto \mstate'}{\mstate \stepto \Gamma(\mstate')}
    \end{align*}
  \end{minipage}
  \begin{minipage}{0.65\linewidth}
    \begin{align*}
      \Gamma(\tpl{(\mexpr,\menv),\mstore,\mkont, \mktab, \mmemo}) &= \tpl{(\mexpr,\menv),\mstore|_{\reaches(\mathit{root}, \mstore)}, \mkont, \mktab, \mmemo} \\
      \text{where } \mathit{root} &= \touchesk{\mkont}(\mktab) \cup
      \mathit{range}(\menv|_{\mathit{fv}(\mexpr)})
    \end{align*}
  \end{minipage}
\end{center}
%
The ``reaches'' metafunction ($\reaches$) is a transitive closure of ``touches'' as it weaves through the store, starting from some root set of addresses:

@@

The creators of CFA2 had a clear goal of harnessing the extra information a pushdown model provides to produce a high-precision analysis that works well in practice.
%
This resulted in the following features in addition to just the call/return matching of \autoref{sec:pdcfa}:
%

@@

%%
The second of these is to ameliorate a problem they call ``fake rebinding.''
%
Bindings in the abstract represent several values, but we don't want to reference a variable \texttt{x} in different places in a function and have it resolve to different values
\footnote{In AAM, a variable reference non-deterministically steps to all possible values associated with that variable.}.
%
If subsequent variable references are not the same, it looks as if the variable was rebound; it hasn't, and thus it is a ``fake rebinding.''
%
%CFA2 does not step to all values on variable reference, but instead carries all its values around in superposition until they need to be observed at, say, a function call.
%
%We give a simplified semantics that is more along the AAM style, but CFA2's approach can easily be recovered from it.\footnote{Our Redex model implements the fake rebinding strategy CFA2 itself employs.}
%%

@@

A tabular semantics for the CES$\xi$K machine follows the previous section, except that $\mktab$ now stores $(\mkont, \mframe)$ pairs, and returns reinstate the saved $\mframe$:
\begin{align*}
  \tpl{\mval,\mstore,\mframe,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont},\mktab,\mmemo} &\stepto
  \tpl{\mpoint,
    \mstore',
    \mframe',
    \krt{\mpoint, \mstore'},
    \mktab',
    \mmemo} \\
  \text{where }
  & \mpoint = (\mexpr, \extm{\menv}{\mvar}{\maddr}) \\
  & (\mstore',\mframe') = \bind(\mstore,\maddr,\mvar,\mval) \\
  & \mktab' = \joinone{\mktab}{(\mpoint, \mstore')}{(\mkont,\mframe)}
\\[2pt]
 \tpl{\mval, \mstore, \mframe, \krt{\mpoint,\mstore'}, \mktab, \mmemo} &\stepto
 \tpl{\mval, \mstore, \mframe', \mkont, \mktab, \joinone{\mmemo}{(\mpoint, \mstore')}{(\mval,\mstore)}}
  \\ & \quad\text{if } (\mkont,\mframe') \in \mktab(\mpoint, \mstore')
\end{align*}

@@

\begin{center}
  \begin{minipage}{0.45\linewidth}
    \begin{align*}
      \mstate \in \State &::= \tpl{\mpoint, \mstore, \mkont, \mmkont} \\
      \mpoint \in \Point &::= (\mexpr, \menv) \alt \mval
    \end{align*}
  \end{minipage}
  \begin{minipage}{0.50\linewidth}
    \begin{align*}
      \mval \in \Value &::= \vclo{\slam{\mvar}{\mexpr}}{\menv} \alt \vcomp{\mkont} \\
      \mmkont \in \MKont &::= \kmt \alt \mkapp{\mkont}{\mmkont}
    \end{align*}
  \end{minipage}
\end{center}
\begin{figure}
  \centering
  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
  \begin{tabular}{r|l}%{r|ll}
    \hline
% Reset
    $\tpl{(\sreset{\mexpr}, \menv), \mstore, \mkont, \mmkont}$
    &
    $\tpl{(\mexpr, \menv), \mstore, \kmt, \mkapp{\mkont}{\mmkont}}$
%    & \textsc{[push prompt]}
    \\
% Pop prompt
    $\tpl{\mval, \mstore, \kmt, \mkapp{\mkont}{\mmkont}}$
    &
    $\tpl{\mval, \mstore, {\mkont}, {\mmkont}}$
%    & \textsc{[pop prompt]}
    \\
% Shift
    $\tpl{(\sshift{\mvar}{\mexpr}, \menv), \mstore, \mkont, \mmkont}$
    &
    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \mstore',\kmt,\mmkont}$
%    & \textsc{[capture continuation]}
    \\ & where $\mstore' = \joinone{\mstore}{\maddr}{\vcomp{\mkont}}$
    \\
% continuation call
    $\tpl{\mval,\mstore,\kcons{\kfn{\vcomp{\mkont'}}}{\mkont}, \mmkont}$
    &
    $\tpl{\mval, \mstore, \mkont', \mkapp{\mkont}{\mmkont}}$
%    & \textsc{[compose continuation]}
  \end{tabular}  
  \caption{Machine semantics for shift/reset}
  \label{fig:shift-reset}
\end{figure}

@@

\begin{figure*}
  \centering
  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
  \begin{tabular}{r|l}
    \hline
% Reset
    $\tpl{(\sreset{\mexpr}, \menv), \mstore, \mkont, \mmkont, \mktab, \mmemo}$
    &
    $\tpl{\mpoint, \mstore, \kmt, \kprompt{\mctx}, \joinone{\mktab}{\mctx}{(\mkont,\mmkont)}, \mmemo}$
    \\
    where & $\mpoint = (\mexpr, \menv)$, $\mctx = (\mpoint, \mstore)$.
    \\
% Pop prompt
    $\tpl{\mval, \mstore, \kmt, \kprompt{\mctx}, \mktab, \mmemo}$
    &
    $\tpl{\mval, \mstore, {\mkont}, {\mmkont}, \mktab, \joinone{\mmemo}{\mctx}{(\mval,\mstore)}}$
    if $(\mkont,\mmkont) \in \mktab(\mctx)$
    \\
% Shift
    $\tpl{(\sshift{\mvar}{\mexpr}, \menv), \mstore, \mkont, \mmkont, \mktab, \mmemo}$
    &
    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \joinone{\mstore}{\maddr}{\vcomp{\mkont}},\kmt,\mmkont,\mktab,\mmemo}$
    \\
% continuation call
    $\tpl{\mval,\mstore,\kcons{\kfn{\vcomp{\mkont'}}}{\mkont}, \mmkont, \mktab, \mmemo}$
    &
    $\tpl{\mval, \mstore, \mkont', \kprompt{\mctx}, \mktab',\mmemo}$
    \\ & \quad if $\mctx \notin \dom(\mmemo)$
    \\
    & or \\
    & $\tpl{\mval', \mstore', \mkont, \mmkont, \mktab',\mmemo}$
    \\ & \quad if $(\mval',\mstore') \in \mmemo(\mctx)$
    \\
    where & $\mctx = (\vcomp{\mkont'}, \mval, \mstore)$ \\
          & $\mktab' = \joinone{\mktab}{\mctx}{(\mkont,\mmkont)}$
    \\
% Memoized continuation call
% continuation call
    $\tpl{\mval,\mstore,\kcons{\kfn{\vcomp{\mkont'}}}{\mkont}, \mmkont, \mktab, \mmemo}$
    &
    $\tpl{\mval', \mstore', \mkont, \mmkont, \mktab,\mmemo}$ if $(\mval',\mstore') \in \mmemo(\vcomp{\mkont'},\mval,\mstore)$
  \end{tabular}  
  \caption{Faulty table-based semantics for shift/reset}
  \label{fig:shift-reset-table0}
\end{figure*}
